Title Goes Here

### 2019-10-28 analysis/paper/paper.Rmd
revised submission

author: 
  - name          : "Daniel Pearson"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "School of Psychology, University of New South Wales, UNSW Sydney NSW 2052, Australia"
    email         : "d.pearson@unsw.edu.au"
  - name          : "Poppy Watson"
    affiliation   : "1"
  - name          : "Phillip (Xin) Cheng"
    affiliation   : "2"
  - name          : "Mike E. Le Pelley"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "School of Psychology, UNSW Sydney"
  - id            : "2"
    institution   : "Department of Cognitive Science, Macquarie University"

@@

abstract: |
  Salient-but-irrelevant distractors can automatically capture attention and eye-gaze in visual search. However, recent findings have suggested that attention to salient-but-irrelevant stimuli can be suppressed when observers use a specific target template to guide their search (i.e., feature search). Increasingly, evidence suggests that attentional selection is influenced by factors other than the physical salience of a stimulus and the observer’s goals. For instance, pairing a stimulus with reward has been shown to increase the extent to which it captures attention and gaze (as though it has become more physically salient), even when such capture has negative consequences for the observer. We investigated whether capture by reward can be suppressed in the same way as capture by physical salience using eye-tracking with a rewarded visual search task. When participants were encouraged to use feature search, attention to a distractor paired with relatively small reward was suppressed. By contrast, attention was captured by a distractor paired with large reward, even when such capture resulted in the loss of large reward. These findings suggest that reward-related stimuli are given special priority to the visual attention system and have implications for our understanding of real-world biases to reward-related stimuli, such as those seen in addiction.

keywords          : "attentional capture, reward, suppression, cognitive control"
wordcount         : "6787"

@@

For example, across two experiments @Gaspelin2016 had participants perform variants of a commonly used  procedure for examining attentional capture---the *additional singleton task* [@Theeuwes1991a; @Theeuwes1992]---while recording eye movements. Eye movements provide a useful index of attention since an eye movement to a location (a saccade) is necessarily preceded by a shift of spatial attention to that location [@Deubel1996]. In this task, participants were required to locate and respond to a shape-singleton target that was presented in an array of non-target shapes. On half of the trials, one of the non-target shapes was rendered in a different colour from the rest of the array so as to be a physically salient colour-singleton distractor. If participants were *more* likely to make a saccade towards the singleton distractor than one of the nonsingleton non-targets, this would indicate that the singleton distractor had captured attention. By contrast, if participants were *less* likely to make a saccade to the singleton distractor than one of the nonsingleton non-targets, this would suggest that attention to the singleton distractor had been suppressed. In @Gaspelin2016\'s Experiment 1, participants were encouraged to adopt a strategy of searching for the "odd-one-out" in the display [i.e., *singleton-search mode*; @Bacon1994], by having the target be randomly either a circle amongst diamonds or a diamond amongst circles on each trial (see Figure \@ref(fig:methodFig)A for an example from the current study). Under these conditions, both stimulus-driven [@Theeuwes1992] and goal-directed [@Bacon1994] accounts of attentional selection would predict capture, as the template that participants use to locate the target (i.e., any singleton) is broad enough to include the singleton distractor. As expected, under these singleton-search conditions Gaspelin et al. found that participants' overt attention was captured by the salient singleton. In Experiment 2, participants were instead encouraged to adopt a more specific target template [i.e., *feature-search mode*; @Bacon1994], by consistently mapping the target identity to one shape (e.g., search for a circle on every trial), and having a heterogeneous set of shapes act as the non-targets (see Figure \@ref(fig:methodFig)B). This manipulation makes searching for any singleton a highly inefficient strategy, as there are multiple singleton nontargets in the display (e.g., a circle, a square, a triangle, and a red shape). Thus, participants should be encouraged to search for the specific features that define the target (i.e., the "circleness" of the circle). Under these conditions, attention to the salient colour-singleton distractor was suppressed relative to the other non-targets, suggesting that the visual attention system treated the salient singleton differently from the other non-targets (inconsistent with goal-directed accounts), yet did not capture gaze (inconsistent with stimulus-driven accounts). By contrast, the signal suppression hypothesis can account for these findings by suggesting that the singleton generated an attentional priority signal as a consequence of its physical salience, but this signal could be suppressed by an inhibitory mechanism under conditions that promoted a more specific search strategy (i.e., feature search). Importantly, this suppression effect was observed even for the most rapidly generated eye movements, which was taken as evidence that the suppression effect operates early in processing, before attention is initially allocated.

## The role of selection history in guiding attention
Up to this point we have discussed the set of influences on attentional control as a dichotomy: goal-directed versus stimulus-driven. However, a growing literature has questioned this framework, pointing to demonstrations of attentional selection biases that fall outside this dichotomy [@Awh2012; @LePelleya; @Failing2018]. @Awh2012 suggested a third category of influences on attentional selection, labelled *selection history*. This category describes attentional biases that are formed through learned relationships between stimulus features and motivationally significant outcomes---such as reward [@Anderson2015a; @Failing2018; @LePelleya] or punishment [@Watson2019]---as well as other biases related to the history of attentional deployment that cannot be easily explained in terms of goal-directed or stimulus-driven control, such as biases for selecting stimulus features that have previously acted as targets [e.g., @Maljkovic1994].

The influence of learned stimulus-reward associations on attentional capture is demonstrated in a study by Le Pelley et al. [-@LePelley2015b; see also @Pearson2015]. In a variant of the additional singleton task, participants were more likely to make a saccade towards a colour-singleton distractor signalling the availability of high reward (e.g., 10$\text{\textcent{}}$) than a singleton distractor signalling low reward (1$\text{\textcent{}}$), even though looking at the reward-signalling singletons was counterproductive, as it resulted in the omission of the reward that would otherwise have been delivered. That is, participants were more likely to have their attention and gaze captured by a distractor signalling high reward, even though this led to the omission of more high rewards. This result cannot be explained in terms of stimulus-driven influences, as the colour-reward associations were counterbalanced across participants to avoid any influence of differential physical salience on capture. Similarly, the effect cannot be explained by goal-directed influences, as capture by the reward associated distractor was directly counterproductive to participants' goal of maximizing reward. Thus, this study demonstrates that experiencing the relationship between a stimulus feature and reward influences the likelihood that that feature will capture attention and gaze, a finding termed *value-modulated attentional capture* (VMAC).

Pairing a stimulus with reward is thought to increase its attentional priority in a way that mimics or otherwise interacts with its physical salience. For example, eye-tracking studies have demonstrated that VMAC is largest for the most rapidly generated saccades, and diminishes with increasing saccade latency [@Pearson2016; @Failing2015], a pattern similar to that observed in capture by physical salience [e.g., @VanZoest2004]. Similarly, electrophysiological studies have shown that reward affects ERP components thought to index low-level perceptual activity [@Hickey2010]. This is consistent with the idea of *incentive salience* [@Berridge1998], wherein a reward-related stimulus becomes more salient and attention grabbing to the observer through a change in its perceptual representation. This raises the question, if associating a stimulus with reward increases its perceptual priority as though it has become more physically salient, can this priority signal be suppressed by the inhibitory control process proposed by the signal suppression hypothesis? That is, can incentive salience be suppressed in the same way as physical salience? Aside from the theoretical importance of this question, it may also have important clinical implications. Recent evidence suggests that the processes underlying VMAC are closely related to the attentional biases towards drug-related stimuli observed in addiction [@Albertella2017a; @Anderson2013; for review see @Field2008], which can promote relapse in recovering addicts [@Marissen2006; @Waters2012; @Marhe2013]. Thus it is important to know whether cognitive control processes can be used to overcome the influence of reward on attention when such capture is counterproductive and contrary to our goals. We addressed this question by combining Le Pelley et al.'s VMAC procedure with Gaspelin et al.'s (2017) overt attentional suppression procedure to investigate whether reward-related attentional biases can be suppressed by inhibitory control processes under conditions promoting feature search.

@@

## Stimuli
The task was based on that used by @Gaspelin2016 and Pearson et al. [-@Pearson2015; -@Pearson2016]. Each trial consisted of a fixation display, a search display, and a feedback display (see Figure \@ref(fig:methodFig)C). All stimuli were presented on a black background. The fixation display consisted of a white cross (0.5$^\circ$ visual angle), presented in the centre of the screen and surrounded by a white circle (diameter 3.0$^\circ$). The search display consisted of four filled shapes (size 1.6$^\circ$ $\times$ 1.6$^\circ$) presented above, below, left and right of screen centre (eccentricity 4.5$^\circ$). One of the shapes was designated as the target, and the other three shapes were designated non-targets. For all participants, the target was randomly determined to be either a circle or a diamond on each trial [^1]. The identity of the non-target shapes varied by search condition. For the *singleton search* condition, the non-target stimuli were all rendered as the alternative target shape (i.e., if the target was a diamond, all of the non-targets were circles; Figure \@ref(fig:methodFig)A). For the *feature search* condition, the non-target stimuli were randomly selected without replacement from a set of two squares and two hexagons (such that one of the non-targets was always a shape singleton), and were never rendered as the alternative target shape (Figure \@ref(fig:methodFig)B). The target and two of the non-targets were always rendered in grey (CIE $x$, $y$ chromaticity coordinates of .327/.400). On the majority of trials, one of the non-target shapes (the *singleton distractor*) was rendered in either orange or blue (CIE $x$, $y$ coordinates .492/.445 and .192/.216, respectively). On the remaining trials, all shapes were rendered in grey such that there was no singleton distractor. The values of blue and orange had similar luminance (~24.5 cd/m^2^), which was higher than that of grey (~8.3 cd/m^2^). The feedback display showed the points earned on the current trial. If the response time (RT) exceeded the soft-timeout threshold (see below), the message "Too slow" appeared below the reward feedback.

@@

On each trial, the location of the target was determined randomly. On high-reward and low-reward trials, the location of the singleton distractor was randomly selected from the three remaining locations. A small, circular region of interest (ROI) with diameter 1.7$^\circ$ visual angle was defined around the centre of the target, with a larger ROI (diameter 2.6$^\circ$) defined around the singleton distractor. A response was registered when the participant had accumulated 100 ms of gaze dwell time within the target ROI. If any gaze was detected within the singleton ROI, the reward for the trial was not awarded (henceforth referred to as *reward omissions*). On distractor-absent trials, one of the non-target shapes was selected at random to act as the "singleton distractor" location. Any gaze falling within the ROI around this (non-salient) stimulus triggered a reward omission in the same way as on trials with a singleton distractor. Responses with RTs that were slower than the soft-timeout threshold (1000 ms) were also not rewarded.

## Procedure
Participants were told that their task was to move their eyes to the target "as quickly and directly as possible" on each trial. Those in the singleton search condition were instructed to move their eyes to the "unique shape" on each trial, whereas those in the feature search condition were instructed to move their eyes to the "circle or diamond". Participants were also informed about the relationships between distractor colour and reward availability (e.g., that if an orange shape was present in the display they could 500 points, and if a blue shape was present they could win 10 points); they were not informed that looking at the reward-associated distractor would cancel the reward for the current trial. The instructions stated that they would earn 0 points if their RT was slower than the soft-timeout threshold, and that their reward also depended on how accurately they moved their eyes to the target. Participants were told that they could earn between 7 and 15 AUD for good performance, but no specific information about the conversion rate from points to AUD was provided.

@@

## Data analysis
In line with previous VMAC protocols [e.g., @LePelley2015b; @Pearson2016], data from the first two experimental trials, and the first two trials after each break were discarded. Hard timeouts (`r op(hardtimeouts$perc[hardtimeouts$timeout == 1], 1)`\% of all trials) were also discarded. One participant was replaced because their mean proportion of valid gaze samples during each trial was less than 50\%. One other participant was replaced after reporting that they had misunderstood the task instructions. For the remaining participants, valid gaze data were registered on an average of `r op(allTracks_summ$mean_trial_prop, 1)`\% ($SD=$ `r op(allTracks_summ$sd_trial_prop, 1)`\%) of samples, suggesting high fidelity of gaze data.

Saccades were detected in the raw gaze data (sampled at 300 Hz) using a velocity-threshold identification algorithm [@Salvucci2000]. Saccade latency was measured as the duration from onset of the search display to the point at which eye-movement velocity exceeded 40$^\circ$/s. Following previous protocols [@Pearson2016], we excluded trials with anticipatory saccades (saccade latency <80 ms; `r op(totalSaccadeTrialsDiscarded$prop_anticipatory,1)`\% of all trials), trials in which no gaze was recorded within 100 pixels (5.1$^\circ$) of the fixation point within the first 80 ms (`r op(totalSaccadeTrialsDiscarded$prop_outside_fix, 1)`\% of trials), and trials in which there was insufficient gaze data to detect a saccade (`r op(totalSaccadeTrialsDiscarded$prop_bad_data, 1)`\% of trials). To classify the direction of the first saccade on each trial, the angular deviation between the endpoint of the saccade and each stimulus location was calculated. The saccade was defined as going in the direction of a stimulus if the angular deviation was less than $30^\circ$ to the left or right of that stimulus.

There were two primary dependent measures: (1) the percentage of reward omissions---that is, trials on which participants looked at the singleton distractor (or nonsingleton distractor randomly assigned to act as the "singleton distractor" location: see Design) and hence reward was cancelled---for high-reward, low-reward, and distractor-absent trials, and (2) the percentage of first saccades directed towards the singleton distractor versus the average nonsingleton distractor (that is, the mean across the two nonsingleton distractors present in each search display; see Figure \@ref(fig:methodFig)), on high-reward and low-reward trials.

For the analysis of reward omissions, we were particularly interested in two contrasts. First, the comparison between the high-reward and low-reward trials allowed us to examine the effect of reward on attention, since both displays contained a physically salient colour-singleton distractor, with the only difference being the magnitude of reward that the distractor signalled. Second, the comparison between the low-reward and distractor-absent trials allowed us to examine the effect of physical salience on attention, since these different trial types differed in whether the display contained a colour-singleton distractor, but had the same reward magnitude available (10 points). These contrasts were compared for the singleton search and feature search conditions using 2 $\times$ 2 mixed ANOVAs.

For the analysis of saccade direction, we were primarily interested in the difference between the percentage of trials on which first saccades were directed towards the singleton distractor and the mean of both nonsingleton distractors across high- and low-reward trials. These data were converted into a *capture score* by subtracting the percentage of trials where the first saccade was directed towards the average nonsingleton from the percentage of trials where the first saccade was directed towards the singleton. A positive capture score indicated that gaze was more likely to have been directed towards the singleton than the average nonsingleton (i.e., the singleton had captured overt attention), whereas a negative capture score indicated that gaze was more likely to be directed toward the average nonsingleton than the singleton (overt attention to the singleton was suppressed). A 2 $\times$ 2 mixed ANOVA was used to compare capture scores for high- and low-reward trials across search conditions. In addition to any group differences in the patterns of attentional capture and/or suppression across trial type, we were also interested in directly investigating whether attention was captured by the singleton distractor (capture score $>0$), or whether attention to the singleton was suppressed (capture score $<0$) for each trial type $\times$ search condition combination. We therefore tested the capture score for each trial type and search condition with planned one-sample *t*-tests against zero.

Statistical analyses were conducted in `r cite_r()` with the afex package  [@R-afex] used to calculate ANOVAs and the emmeans package [@R-emmeans] for follow-up contrasts. Greenhouse-Geisser corrections are reported where appropriate. In cases where a conclusion is drawn on the basis of a null effect, we report the Bayes factor that corresponds to a Bayesian *t*-test using the default Cauchy prior, conducted using the BayesFactor package [@R-BayesFactor]. Bayes
factors are interpreted in line with the guidelines suggested by @Jeffreys1961. Experiment scripts, raw data, and a reproducible research compendium of analysis scripts [@R-rrtools] are available at https://osf.io/yrdzv/.

@@

### Effect of physical salience.
ANOVA analysis of the percentage of reward omissions revealed a non-significant main effect of trial type (low-reward vs distractor-absent), `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$dType`, and a significant main effect of search condition (singleton search vs feature search), `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$condition`. Critically, this main effect was qualified by a significant trial type $\times$ search condition interaction, `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$condition_dType`, indicating that the effect of physical salience on overt attentional capture differed across search conditions. Paired-samples *t*-tests revealed that participants in the singleton search condition triggered significantly more reward omissions on trials that contained a colour-singleton distractor than those that did not, `r apa_print(singleton_physicalsalience_omissions.ttest)$statistic`, $d_z=$ `r singleton_physicalsalience_omissions.es`, suggesting that overt attention was captured by the physically salient singleton. In contrast, participants in the feature search condition triggered fewer reward omissions on trials containing a colour-singleton distractor than trials with no coloured singleton, `r apa_print(feature_physicalsalience_omissions.ttest)$statistic`, $d_z=$ `r feature_physicalsalience_omissions.es`, suggesting that overt attention to the singleton distractor was suppressed under conditions promoting feature search, consistent with previous findings [@Gaspelin2016].

>>>>

author: 
  - name          : "Daniel Pearson"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Institute of Cognitive Neuroscience, University College London, Gower Street, London WC1E 6BT, UK"
    email         : "daniel.pearson@ucl.ac.uk"
  - name          : "Poppy Watson"
    affiliation   : "2"
  - name          : "Phillip (Xin) Cheng"
    affiliation   : "3"
  - name          : "Mike E. Le Pelley"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Institute of Cognitive Neuroscience, University College London"
  - id            : "2"
    institution   : "School of Psychology, UNSW Sydney"
  - id            : "3"
    institution   : "Department of Cognitive Science, Macquarie University"

@@

abstract: |
  Salient-but-irrelevant distractors can automatically capture attention and eye-gaze in visual search. However, recent findings have suggested that attention to salient-but-irrelevant stimuli can be suppressed when observers use a specific target template to guide their search (i.e., feature search). A separate line of research has indicated that attentional selection is influenced by factors other than the physical salience of a stimulus and the observer’s goals. For instance, pairing a stimulus with reward has been shown to increase the extent to which it captures attention and gaze (as though it has become more physically salient), even when such capture has negative consequences for the observer. Here we used eye-tracking with a rewarded visual search task to investigate whether capture by reward can be suppressed in the same way as capture by physical salience. When participants were encouraged to use feature search, attention to a distractor paired with relatively small reward was suppressed. However, under the same conditions attention was captured by a distractor paired with large reward, even when such capture resulted in reward omission. These findings suggest that reward-related stimuli are given special priority within the visual attention system over and above physically-salient stimuli, and have implications for our understanding of real-world biases to reward-related stimuli, such as those seen in addiction.

keywords          : "attentional capture, reward, suppression, cognitive control"
wordcount         : "8515"

@@

For example, across two experiments @Gaspelin2016 had participants perform variants of a commonly used  procedure for examining attentional capture---the *additional singleton task* [@Theeuwes1991a; @Theeuwes1992]---while recording eye movements. Eye movements provide a useful index of attention since an eye movement to a location (a saccade) is necessarily preceded by a shift of spatial attention to that location [@Deubel1996]. In this task, participants were required to locate and respond to a shape-singleton target that was presented in an array of non-target shapes. On half of the trials, one of the non-target shapes was rendered in a different colour from the rest of the array so as to be a physically salient colour-singleton distractor. If participants were *more* likely to make a saccade towards the singleton distractor than one of the nonsingleton non-targets, this would indicate that the singleton distractor had captured attention. By contrast, if participants were *less* likely to make a saccade to the singleton distractor than one of the nonsingleton non-targets, this would suggest that attention to the singleton distractor had been suppressed. In Gaspelin et al.'s Experiment 1, participants were encouraged to adopt a strategy of searching for the "odd-one-out" in the display [i.e., *singleton-search mode*; @Bacon1994], by having the target be randomly either a circle amongst diamonds or a diamond amongst circles on each trial (see Figure \@ref(fig:methodFig)A for an example from the current study). Under these conditions, both stimulus-driven [@Theeuwes1992] and goal-directed [@Bacon1994] accounts of attentional selection would predict capture, as the template that participants use to locate the target (i.e., any singleton) is broad enough to include the singleton distractor. As expected, under these singleton-search conditions Gaspelin et al. found that participants' overt attention was captured by the salient singleton. In Experiment 2, participants were instead encouraged to adopt a more specific target template [i.e., *feature-search mode*; @Bacon1994], by consistently mapping the target identity to one shape (e.g., search for a circle on every trial), and having a heterogeneous set of shapes act as the non-targets (see Figure \@ref(fig:methodFig)B). This manipulation makes searching for any singleton a highly inefficient strategy, as there are multiple singleton non-targets in the display (e.g., a circle, a square, a triangle, and a red shape). Thus, participants should be encouraged to search for the specific features that define the target (i.e., the "circleness" of the circle). Under these conditions, attention to the salient colour-singleton distractor was suppressed relative to the other non-targets, suggesting that the visual attention system treated the salient singleton differently from the other non-targets (inconsistent with goal-directed accounts), yet did not capture gaze (inconsistent with stimulus-driven accounts). By contrast, the signal suppression hypothesis can account for these findings by suggesting that the singleton generated an attentional priority signal as a consequence of its physical salience, but this signal could be suppressed by an inhibitory mechanism under conditions that promoted a more specific search strategy (i.e., feature search). Importantly, this suppression effect was observed even for the most rapidly generated eye movements, which was taken as evidence that the suppression effect operates early in processing, before attention is initially allocated.

## The role of selection history in guiding attention
Up to this point we have discussed the set of influences on attentional control as either goal-directed or stimulus-driven. However, a growing literature has questioned this framework, pointing to demonstrations of attentional selection biases that fall outside this dichotomy [@Awh2012; @LePelleya; @Failing2018]. @Awh2012 suggested a third category of influences on attentional selection, labelled *selection history*. This category describes attentional biases that are formed through learned relationships between stimulus features and motivationally significant outcomes---such as reward [@Anderson2015a; @Failing2018; @LePelleya] or punishment [@Watson2019]---as well as other biases related to the history of attentional deployment that cannot be easily explained in terms of goal-directed or stimulus-driven control, such as biases for selecting stimulus features that have previously acted as targets [e.g., @Maljkovic1994].

The influence of learned stimulus-reward associations on attentional capture is demonstrated in a study by Le Pelley et al. [-@LePelley2015b; see also @Pearson2015]. In a variant of the additional singleton task, participants were more likely to make a saccade towards a colour-singleton distractor signalling the availability of high reward (e.g., 10¢) than a singleton distractor signalling low reward (1¢), even though looking at the reward-signalling singletons was counterproductive, as it resulted in the omission of the reward that would otherwise have been delivered. That is, participants were more likely to have their attention and gaze captured by a distractor signalling high reward, even though this led to the omission of more high rewards. This result cannot be explained in terms of stimulus-driven influences, as the colour-reward associations were counterbalanced across participants to avoid any influence of differential physical salience on capture. Similarly, the effect cannot be explained by goal-directed influences, as capture by the reward associated distractor was directly counterproductive to participants' goal of maximizing reward. Thus, this study demonstrates that experiencing the relationship between a stimulus feature and reward influences the likelihood that that feature will capture attention and gaze, a finding termed *value-modulated attentional capture* (VMAC).

Pairing a stimulus with reward is thought to increase its attentional priority in a way that mimics or otherwise interacts with its physical salience. For example, eye-tracking studies have demonstrated that VMAC is largest for the most rapidly generated saccades, and diminishes with increasing saccade latency [@Pearson2016; @Failing2015], a pattern similar to that observed in capture by physical salience [e.g., @VanZoest2004]. Similarly, neuroimaging and electrophysiological studies have shown that the influence of reward arises early on in visual processing, with reward effects observed as early as low-level visual cortex [e.g., @MacLean2015; @Serences2008; @Serences2010]. Considered together, these findings suggest that the attentional priority of a reward-related stimulus is increased (at least in part) through the augmentation of its stimulus-driven perceptual activity, akin to increasing its *incentive salience* [@Berridge1998]. This raises the question, if associating a stimulus with reward increases its perceptual priority as though it has become more physically salient, can this priority signal be suppressed by the inhibitory control process proposed by the signal suppression hypothesis? That is, can incentive salience be suppressed in the same way as physical salience? This question formed the focus of the current study. Aside from the theoretical importance of this question, it may also have important clinical implications. Recent evidence suggests that the processes underlying VMAC are closely related to the attentional biases towards drug-related stimuli observed in addiction [@Albertella2017a; @Albertella2019; @Albertella2019a; @Anderson2013; for review see @Field2008], which can promote relapse in recovering addicts [@Marissen2006; @Waters2012; @Marhe2013]. Thus it is important to know whether cognitive control processes can be used to overcome the influence of reward on attention when such capture is counterproductive and contrary to our goals. We addressed this question by combining Le Pelley et al.'s VMAC procedure with Gaspelin et al.'s (2017) overt attentional suppression procedure to investigate whether reward-related attentional biases can be suppressed by inhibitory control processes under conditions promoting feature search.

@@

## Stimuli
The task was based on that used by @Gaspelin2016 and Pearson et al. [-@Pearson2015; -@Pearson2016]. Each trial consisted of a fixation display, a search display, and a feedback display (see Figure \@ref(fig:methodFig)C). All stimuli were presented on a black background. The fixation display consisted of a white cross ($0.5^\circ$ visual angle), presented in the centre of the screen and surrounded by a white circle (diameter $3.0^\circ$). The search display consisted of four filled shapes (size $1.6^\circ \times 1.6^\circ$) presented above, below, left and right of screen centre (eccentricity $4.5^\circ$). One of the shapes was designated as the target, and the other three shapes were designated non-targets. For all participants, the target was randomly determined to be either a circle or a diamond on each trial [^1]. The identity of the non-target shapes varied by search condition. For the *singleton search* condition, the non-target stimuli were all rendered as the alternative target shape (i.e., if the target was a diamond, all of the non-targets were circles; Figure \@ref(fig:methodFig)A). For the *feature search* condition, the non-target stimuli were randomly selected without replacement from a set of two squares and two hexagons (such that one of the non-targets was always a shape singleton), and were never rendered as the alternative target shape (Figure \@ref(fig:methodFig)B). The target and two of the non-targets were always rendered in grey (CIE $x$, $y$ chromaticity coordinates of .327/.400). On the majority of trials, one of the non-target shapes (the *singleton distractor*) was rendered in either orange or blue (CIE $x$, $y$ coordinates .492/.445 and .192/.216, respectively). On the remaining trials, all shapes were rendered in grey such that there was no singleton distractor. The values of blue and orange had similar luminance (~24.5 cd/m^2^), which was higher than that of grey (~8.3 cd/m^2^). The feedback display showed the points earned on the current trial. If the response time (RT) exceeded 1000 ms, the message "Too slow" appeared below the reward feedback.

@@

On each trial, the location of the target was determined randomly. On high-reward and low-reward trials, the location of the singleton distractor was randomly selected from the three remaining locations. A small, circular region of interest (ROI) with diameter $1.7^\circ$ visual angle was defined around the centre of the target, with a larger ROI (diameter $2.6^\circ$) defined around the singleton distractor. A response was registered when the participant had accumulated 100 ms of gaze dwell time within the target ROI. If any gaze was detected within the singleton ROI, the reward for the trial was not awarded (henceforth referred to as *reward omissions*). On distractor-absent trials, one of the non-target shapes was selected at random to act as the "singleton distractor" location. Any gaze falling within the ROI around this (non-salient) stimulus triggered a reward omission in the same way as on trials with a singleton distractor. Responses with RTs that were slower than 1000 ms were also not rewarded.

## Procedure
Participants were told that their task was to move their eyes to the target "as quickly and directly as possible" on each trial. Those in the singleton search condition were instructed to move their eyes to the "unique shape" on each trial, whereas those in the feature search condition were instructed to move their eyes to the "circle or diamond". Participants were also informed about the relationships between distractor colour and reward availability (e.g., that if an orange shape was present in the display they could win 500 points, and if a blue shape was present they could win 10 points); they were not informed that looking at the reward-associated distractor would cancel the reward for the current trial [previous research suggests that whether or not participants are explicitly informed of this has no influence on the magnitude of the VMAC effect that is observed: @Pearson2015]. The instructions stated that participants would earn 0 points if their RT was slower than 1000 ms, and that their reward also depended on how accurately they moved their eyes to the target. Participants were told that they could earn between 7 and 15 AUD for good performance, but no specific information about the conversion rate from points to AUD was provided.

@@

Consistent with previous VMAC protocols [@LePelley2015b; @Pearson2015; @Pearson2016], participants completed a brief questionnaire after finishing the visual search task, to assess their knowledge of the colour--reward contingencies, and of the omission contingency (i.e., that looking at the coloured distractor caused omission of reward). Participants were presented with an orange and blue circle, in random order, and for each were asked to indicate whether they would earn 0 points, 10 points, or 500 points when the display contained a circle of that colour and they (1) moved their eyes quickly and accurately to the target, or (2) looked at the coloured shape before looking at the diamond. As this study was primarily concerned with the role of non-volitional attentional control processes in attentional suppression of reward-related stimuli [cf. @Gaspelin2018a], we did not have an explicit hypothesis to motivate our analysis of these questionnaire data. However, in the interest of completeness we present an exploratory analysis of these data in supplemental materials.

## Data analysis
In line with previous protocols [e.g., @LePelley2015b; @Pearson2016], data from the first two experimental trials, and the first two trials after each break were discarded. Hard timeouts (`r op(hardtimeouts$perc[hardtimeouts$timeout == 1], 1)`\% of all trials) were also discarded. One participant was replaced because their mean proportion of valid gaze samples during each trial was less than 50\%. One other participant was replaced after reporting that they had misunderstood the task instructions. For the remaining participants, valid gaze data were registered on an average of `r op(allTracks_summ$mean_trial_prop, 1)`\% ($SD=$ `r op(allTracks_summ$sd_trial_prop, 1)`\%) of samples, suggesting high fidelity of gaze data.

Saccades were detected in the raw gaze data (sampled at 300 Hz) using a velocity-threshold identification algorithm [@Salvucci2000]. Saccade latency was measured as the duration from onset of the search display to the point at which eye-movement velocity exceeded $40^\circ$/s. Following previous protocols [@Pearson2016], we excluded trials with anticipatory saccades (saccade latency <80 ms; `r op(totalSaccadeTrialsDiscarded$prop_anticipatory,1)`\% of all trials), trials in which no gaze was recorded within 100 pixels ($5.1^\circ$) of the fixation point within the first 80 ms (`r op(totalSaccadeTrialsDiscarded$prop_outside_fix, 1)`\% of trials), and trials in which there was insufficient gaze data to detect a saccade (`r op(totalSaccadeTrialsDiscarded$prop_bad_data, 1)`\% of trials). To classify the direction of the first saccade on each trial, the angular deviation between the endpoint of the saccade and each stimulus location was calculated. The saccade was defined as going in the direction of a stimulus if the angular deviation was less than $30^\circ$ clockwise or anticlockwise from the centre of that stimulus.

There were two primary dependent measures: (1) the percentage of reward omissions---that is, trials on which participants looked at the singleton distractor (or nonsingleton non-target randomly assigned to act as the "singleton distractor" location; see Design) and hence reward was cancelled---for high-reward, low-reward, and distractor-absent trials, and (2) the percentage of first saccades directed towards the singleton distractor versus the average nonsingleton distractor---that is, the mean across the two nonsingleton non-targets present in each search display; see Figure \@ref(fig:methodFig)---on high-reward and low-reward trials.

For the analysis of reward omissions, we were particularly interested in two contrasts. First, the comparison between the high-reward and low-reward trials allowed us to examine the effect of reward on attention, since both displays contained a physically salient colour-singleton distractor that provided information about the reward available on that trial, with the only difference being the *magnitude* of reward that the distractor signalled. Second, the comparison between the low-reward and distractor-absent trials allowed us to examine the effect of physical salience on attention, since these trial types differed in whether the display contained a colour-singleton distractor, but had the same reward magnitude available (10 points). These contrasts were compared for the singleton search and feature search conditions using 2 $\times$ 2 mixed ANOVAs.

For the analysis of saccade direction, we were primarily interested in the difference between the percentage of trials on which first saccades were directed towards the singleton distractor and the mean of both nonsingleton non-targets across high- and low-reward trials. These data were converted into a *capture score* by subtracting the percentage of trials where the first saccade was directed towards the average nonsingleton from the percentage of trials where the first saccade was directed towards the singleton. A positive capture score indicated that gaze was more likely to have been directed towards the singleton than the average nonsingleton (i.e., the singleton had captured overt attention), whereas a negative capture score indicated that gaze was more likely to be directed toward the average nonsingleton than the singleton (overt attention to the singleton was suppressed). A 2 $\times$ 2 mixed ANOVA was used to compare capture scores for high- and low-reward trials across search conditions. In addition to any group differences in the patterns of attentional capture and/or suppression across trial type, we were also interested in directly investigating whether attention was captured by the singleton distractor (capture score $>0$), or whether attention to the singleton was suppressed (capture score $<0$) for each trial type $\times$ search condition combination. We therefore tested the capture score for each trial type and search condition with planned one-sample *t*-tests against zero.

Statistical analyses were conducted in `r cite_r()` with the afex package  [@R-afex] used to calculate ANOVAs and the emmeans package [@R-emmeans] for follow-up contrasts. Greenhouse-Geisser corrections are reported where appropriate. In cases where a conclusion is drawn on the basis of a null effect, we report the Bayes factor that corresponds to a Bayesian *t*-test using the default Cauchy prior, conducted using the BayesFactor package [@R-BayesFactor]. Bayes factors are interpreted in line with the guidelines suggested by @Jeffreys1961. Experiment scripts (including verbatim task instructions), raw data, and a reproducible research compendium of analysis scripts [@R-rrtools] are available at https://osf.io/yrdzv/.

@@

### Effect of physical salience.
ANOVA analysis of the percentage of reward omissions revealed a non-significant main effect of trial type (low-reward vs distractor-absent), `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$dType`, and a significant main effect of search condition (singleton search vs feature search), `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$condition`. Critically, this main effect was qualified by a significant trial type $\times$ search condition interaction, `r apa_print.afex_aov(omissions_salience.ANOVA, mse = F, es = "pes")$full_result$condition_dType`, indicating that the effect of physical salience on overt attentional capture differed across search conditions. Paired-samples *t*-tests revealed that participants in the singleton search condition triggered significantly more reward omissions on trials that contained a colour-singleton distractor than on trials without a singleton distractor, `r apa_print(singleton_physicalsalience_omissions.ttest)$statistic`, $d_z=$ `r singleton_physicalsalience_omissions.es`, suggesting that overt attention was captured by the physically salient singleton. By contrast, participants in the feature search condition triggered fewer reward omissions on trials containing a colour-singleton distractor than trials with no coloured singleton, `r apa_print(feature_physicalsalience_omissions.ttest)$statistic`, $d_z=$ `r feature_physicalsalience_omissions.es`, suggesting that overt attention to the singleton distractor was suppressed under conditions promoting feature search, consistent with previous findings [@Gaspelin2016].

Orthogonal contrasts used independent *t*-tests to compare the rate of reward omissions between search conditions across each trial type. On low-reward trials, there was no significant difference between the singleton search and feature search groups in the rate of reward omissions, `r apa_print(low_singleton_vs_feature.ttest)$statistic`, $d_s=$ `r low_singleton_vs_feature.es`. However, on distractor-absent trials participants in the feature search condition triggered significantly more reward omissions (i.e., they were more likely to look at the non-target that had been randomly selected to act as the 'distractor' location) than those in the singleton search condition, `r apa_print(absent_singleton_vs_feature.ttest)$statistic`, $d_s=$ `r absent_singleton_vs_feature.es`.
